Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-11 paper/paper.tex
More prose for store deltas and timestamping.

@@

\section{Sacrificing precision for performance}

Each step in the fixpointer for the above techniques requires joining
several (possibly large) stores together. Even with immutable hashes
that utilize sharing, this is memory-intensive. Worse, this
contributes an obscene fraction of the runtime. We reuse a technique
created in Shivers' dissertation~\cite{ianjohnson:Shivers:1991:CFA},
timestamp approximation, that removes the need for joining whole
stores by manipulating a global store imperatively. This loses
precision because the starting store for each state in a machine
configuration can change between stepping all states and introduce
more flows than originally possible. The example in figure
\ref{fig:timestamp} illustrates this loss in precision. {\bf TODO:}
EXPLANATION OF LOSS. In many applications, this is acceptable for the
performance improvement.

\begin{figure}
TODO
\caption{Precision lossy program in timestamp approximation}
\label{fig:timestamp}
\end{figure}

% For David to fill in.
\subsection{Timestamp}
...
\subsection{Pre-allocated global store}
...

@@

(define (step state)
  (match state
    [(ev \(\msto\) e \(\menv\) k) \(e-dispatch\)] ;; dead if compiled.
    [(co \(\msto\) k v)
     (match k
       ['mt (generator (do (\(\msto\)) ([v <- (force \(\msto\) v)])
        (yield (ans \(\msto\) v))))]
       [(ar\(\superscript\mlab\) e \(\menv\))
        (generator (do (\(\msto\)) ()
          (yield (ev \(\msto\) e \(\menv\) (fn v l)))))
       [(fn\(\superscript\mlab\) f)
        (generator (do (\(\msto\)) ([k <- (get-cont \(\msto\) l)])
         (yield (ap \(\msto\) f v k))))]
       [(fi\(\superscript\mlab\) c a \(\menv\))
        (generator
         (do (\(\msto\)) ([k <- (get-cont \(\msto\) l)]
                  [v <- (force \(\msto\) v)])
           (yield (ev \(\msto\) (if v c a) \(\menv\) k))))])]
    [(ap \(\msto\) fun a k)
     (match fun
       [(clos l x e \(\menv\))
        (generator
         (do (\(\msto\)) ([(\(\menv\)* \(\msto\)*) (bind state)])
           (yield (ev \(\msto\)* e \(\menv\)* k))))]
       [(? op? o)
        (do (\(\msto\)) ([k <- (get-cont \(\msto\) l)]
                 [v <- (force \(\msto\) v)]
                 [v (\(\interpdelta\) o (list v))])
          (yield (co \(\msto\) k v)))
       [_ (generator (do (\(\msto\)) () (continue)))]))]))
\end{alltt}
\caption{Parameterized Abstract$^2$ machine for ISWIM}
\label{fig:paam}
\end{figure}

>>>>

Each step in the fixpointer for the above techniques requires joining
several (possibly large) stores together. Even with immutable hashes
that utilize sharing, this is memory-intensive. Worse, this
contributes an obscene fraction of the runtime.

@@

\subsection{Timestamping stores}

The store delta technique removes overhead of joining large stores,
but for determining if a state has been visited before, we still have
to compare entire stores - an expensive computation.  We modify a
technique created in Shivers'
dissertation~\cite{ianjohnson:Shivers:1991:CFA}, timestamp
approximation, that removes the need for joining whole stores by
manipulating a global store imperatively. The original technique loses
precision because the starting store for each state in a machine
configuration can change between stepping all states and introduce
more flows than originally possible. However, by delaying the store
changes until after all steps have completed, we can maintain our
precision. In a variety of experiments, we observed that the overhead
for storing store-deltas was outweighed by the precision keeping flows
in check. That is to say, the analysis was faster {\it and} more precise.

TODO: technical detail

\subsection{Pre-allocating the store}
TODO

@@

(define (step state)
  (match state
    [(ev \(\msto\) e \(\menv\) k) \(e-dispatch\)] ;; dead if compiled.
    [(co \(\msto\) k v)
     (match k
       ['mt (generator (do (\(\msto\)) ([v <- (force \(\msto\) v)])
        (yield (ans \(\msto\) v))))]
       [(ar\(\superscript\mcntr\subscript\mlab\) e \(\menv\) a)
        (generator
         (do (\(\msto\)) ([\(\msto*\) (join \(\msto\) a\(\superscript{f}\) (force \(\msto\) v))])
           (yield (ev \(\msto*\) e \(\menv\) (fn\(\superscript\mcntr\subscript\mlab\) a\(\superscript{f}\) a)))))
       [(fn\(\superscript\mcntr\subscript\mlab\) fa a)
        (generator
         (do (\(\msto\)) ([\(\msto*\) (join \(\msto\) a\(\superscript{a}\) (force \(\msto\) v))]
                  [k <- (get \(\msto\) a)]
                  [f <- (get \(\msto\) fa)])
           (yield (ap\(\superscript\mcntr\subscript\mlab\) \(\msto*\) f a\(\superscript{a}\) k))))]
       [(fi\(\superscript\mcntr\subscript\mlab\) c a \(\menv\))
        (generator
         (do (\(\msto\)) ([k <- (get \(\msto\) l)]
                  [v <- (force \(\msto\) v)])
           (yield (ev \(\msto\) (if v c a) \(\menv\) k))))])]
    [(ap\(\superscript\mcntr\subscript\mlab\) \(\msto\) fun a k)
     (match fun
       [(clos x e \(\menv\))
        (generator
         (do (\(\msto\)) ([(\(\menv\)* \(\msto\)*) (bind state)])
           (yield (ev \(\msto\)* e \(\menv\)* k))))]
       [(? op? o)
        (do (\(\msto\)) ([k <- (get \(\msto\) l)]
                 [v <- (force \(\msto\) v)]
                 [v (\(\interpdelta\) o (list v))])
          (yield (co \(\msto\) k v)))
       [_ (generator (do (\(\msto\)) () (continue)))]))]))
\end{alltt}
\caption{Parameterized Abstract$^2$ machine for ISWIM}
\label{fig:paam}
\end{figure}
